SIMD\slash SIMT-vectorized Sparse Chemical Kinetic Jacobian\slash Thermo-chemical Source Term Evaluation

### 2018-02-22 paper.tex
updates

\subsubsection{Jacobian evaluation}
\label{S:jacobian_results}
\Cref{F:jacobian_perfomance} shows the performance of the sparse and dense Jacobian evaluations in \texttt{pyJac} on the CPU platforms.
In~\cref{F:sparse_vs_dense}, the mean runtime per condition is presented for the shallow-vectorized Intel OpenCL and OpenMP codes on the \avx/ CPU.
The sparse Jacobian evaluations are slower on both Intel OpenCL and OpenMP due to indirect lookup indexing requirements, as discussed in~\cref{S:sparsity}.
Interestingly the shallow-vectorized OpenCL code is less negatively impacted by the indirect lookup; the sparse OpenMP code is \SIrange{2.5}{7}{$\times$} slower than the dense OpenMP evaluation, while the sparse shallow-vectorized OpenCL code is just \SIrange{1.4}{3.3}{$\times$} slower than its dense counterpart.
Indeed, the shallow-vectorized sparse OpenCL code is faster than the dense OpenMP code in all cases.
\Cref{F:sparse_vs_dense_speedup} shows the speedup of the sparse\slash dense shallow-vectorized OpenCL the over the same sparse\slash dense Jacobian format on OpenMP; a speedup of \SIrange{3.7}{4.2}{$\times$} is exhibited for the dense OpenCL code, increasing to \SIrange{6.6}{8.4}{$\times$} for the sparse Jacobian.

\begin{figure}[htbp]
   \centering
  \begin{subfigure}[t]{0.48\linewidth}
      \includegraphics[width=\textwidth]{sparse_vs_dense.pdf}
      \caption{The mean runtime per-condition of sparse and dense Jacobian evaluation for Intel OpenCL\slash OpenMP on the \avx/ machine.}
      \label{F:sparse_vs_dense}
  \end{subfigure}
  \hfill
  \begin{subfigure}[t]{0.48\linewidth}
      \includegraphics[width=\textwidth]{sparse_vs_dense_speedup.pdf}
      \caption{The speedup of the sparse and dense Jacobian evaluation for Intel OpenCL over the sparse\slash dense OpenMP baseline on the \avx/ machine.}
      \label{F:sparse_vs_dense_speedup}
  \end{subfigure}
  \caption{Performance of sparse and dense Jacobian evaluation in \texttt{pyJac}.}
  \label{F:jacobian_perfomance}
\end{figure}

>>>>

\subsubsection{Jacobian evaluation}
\label{S:jacobian_results}
\Cref{F:jacobian_perfomance} shows the performance of the sparse and dense Jacobian evaluations in \texttt{pyJac} on the CPU platforms.
In~\cref{F:sparse_vs_dense}, the mean runtime per condition is presented for the shallow-vectorized Intel OpenCL and OpenMP codes on the \avx/ CPU.
The sparse Jacobian evaluations are slower on both Intel OpenCL and OpenMP due to indirect lookup indexing requirements, as discussed in~\cref{S:sparsity}.
Interestingly the shallow-vectorized OpenCL code is less negatively impacted by the indirect lookup; the sparse OpenMP code is \SIrange{2.5}{7}{$\times$} slower than the dense OpenMP evaluation, while the sparse shallow-vectorized OpenCL code is just \SIrange{1.4}{3.3}{$\times$} slower than its dense counterpart
As a result, the shallow-vectorized sparse OpenCL code is faster than the dense OpenMP code in all cases on the \avx/ machine (\cref{F:sparse_vs_dense}).
\Cref{F:sparse_vs_dense_speedup} shows the speedup of the sparse\slash dense shallow-vectorized OpenCL the over the same sparse\slash dense Jacobian format on OpenMP; the dense OpenCL is \SIrange{3.7}{4.2}{$\times$} faster than the corresponding dense OpenMP code, this speedup increases to \SIrange{6.6}{8.4}{$\times$} for the sparse Jacobian.
On the \sse/ machine,~\cref{F:sparse_vs_dense_sse} shows similar results; the sparse OpenMP code is the slowest in all cases, and the shallow-vectorized OpenCL code is nearly as fast as the dense OpenMP code.
Once again, the sparse OpenCL code is far less negatively impacted by the indirect lookup, and is only \SIrange{1.8}{3.2}{$\times$} slower than its dense counterpart, while the sparse OpenMP code is significantly (\SIrange{2.25}{6.2}{$\times$}) slower than the dense version.
In~\cref{F:sparse_vs_dense_sse_speedup}, the speedup of the sparse and dense shallow-vectorized OpenCL codes are compared to their OpenMP version; the dense OpenCL code is \SIrange{3.8}{4.2}{$\times$} faster, while the sparse shallow-vectorization achieves a speedup of \SIrange{6.6}{9.1}{$\times$}.

\begin{figure}[htbp]
   \centering
  \begin{subfigure}[t]{0.48\linewidth}
      \includegraphics[width=\textwidth]{sparse_vs_dense.pdf}
      \caption{The mean runtime per-condition of sparse and dense Jacobian evaluation for Intel OpenCL\slash OpenMP on the \avx/ machine.}
      \label{F:sparse_vs_dense}
  \end{subfigure}
  \hfill
  \begin{subfigure}[t]{0.48\linewidth}
      \includegraphics[width=\textwidth]{sparse_vs_dense_speedup.pdf}
      \caption{The speedup of the sparse and dense Jacobian evaluation for Intel OpenCL over the sparse\slash dense OpenMP baseline on the \avx/ machine.}
      \label{F:sparse_vs_dense_speedup}
  \end{subfigure}
  \\
  \begin{subfigure}[t]{0.48\linewidth}
      \includegraphics[width=\textwidth]{sparse_vs_dense_sse.pdf}
      \caption{The mean runtime per-condition of sparse and dense Jacobian evaluation for Intel OpenCL\slash OpenMP on the \sse/ machine.}
      \label{F:sparse_vs_dense_sse}
  \end{subfigure}
  \hfill
  \begin{subfigure}[t]{0.48\linewidth}
      \includegraphics[width=\textwidth]{sparse_vs_dense_sse_speedup.pdf}
      \caption{The speedup of the sparse and dense Jacobian evaluation for Intel OpenCL over the sparse\slash dense OpenMP baseline on the \sse/ machine.}
      \label{F:sparse_vs_dense_sse_speedup}
  \end{subfigure}
  \caption{Performance of sparse and dense Jacobian evaluation on the CPU platforms in \texttt{pyJac}.}
  \label{F:jacobian_perfomance}
\end{figure}

In~\cref{F:sparse_jac_scaling}, the parallel scaling efficiency of the sparse shallow-vectorized OpenCL is compared to the sparse OpenMP code.
The plot is a bit difficult to read, as most of the data-points are clustered together however the key takeaway is that the parallel scaling efficiency of the shallow-vectorized OpenCL code is much closer in value to that of the OpenMP code as compared to the source rate evaluations (\cref{F:source_parallel_scaling}).
The \ce{H2}\slash\ce{CO} model has the worst scaling for both codes, ranging from \numrange{0.93}{0.54} and \numrange{0.99}{0.86} for OpenCL and OpenMP respectively on \numrange{2}{16} cores.
As the model size increases, the efficiency of the OpenCL code improves dramatically, reaching \numrange{0.997}{0.84} for the isopentanol model.
It is worth noting that the OpenCL code's scaling is more negatively affected by the switch from a single 12-core CPU (\cref{t:cpus}) to two CPUs (i.e., the 16 core data-point) than the OpenMP code.
In contrast,~\cref{F:dense_jac_scaling} shows scaling for the shallow-vectorized dense Jacobian OpenCL code that is higher than for source-rate evaluation, but not as good as for sparse Jacobian evaluation.

\begin{figure}[htbp]
   \centering
  \begin{subfigure}[t]{0.48\linewidth}
      \includegraphics[width=\textwidth]{sparse_jac_scaling.pdf}
      \caption{Sparse Jacobian evaluation parallel scaling efficiency for Intel OpenCL\slash OpenMP on the \avx/ machine.}
      \label{F:sparse_jac_scaling}
  \end{subfigure}
  \hfill
  \begin{subfigure}[t]{0.48\linewidth}
      \includegraphics[width=\textwidth]{dense_jac_scaling.pdf}
      \caption{Dense Jacobian evaluation parallel scaling efficiency for Intel OpenCL\slash OpenMP on the \avx/ machine.}
      \label{F:dense_jac_scaling}
  \end{subfigure}
  \caption{The parallel scaling efficiencies of sparse and dense Jacobian evaluation for the shallow-vectorized Intel OpenCL and OpenMP codes on the \avx/ CPU.}
\end{figure}
